---
title: "BMI Calculator and Health Visualizer"
author: "Gabriel Demetrios Lafis"
date: "May 28, 2025"
output: 
  ioslides_presentation:
    css: styles.css
    logo: images/bmi_logo.png
    widescreen: true
    smaller: true
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(ggplot2)
```

## BMI Calculator and Health Visualizer

### An Interactive Shiny Application

- Calculates Body Mass Index (BMI) based on user input
- Provides personalized health recommendations
- Visualizes BMI in relation to health standards
- Projects potential BMI changes over time

## How It Works

The application uses the standard BMI formula:

$$BMI = \frac{weight (kg)}{height (m)^2}$$

```{r, echo=TRUE}
# Sample BMI calculation
height_m <- 1.75  # height in meters
weight_kg <- 70   # weight in kilograms
bmi <- weight_kg / (height_m^2)
cat("BMI for a person who is 1.75m tall and weighs 70kg:", round(bmi, 1))
```

## Interactive Features

The application includes:

1. **User Input Widgets**
   - Height and weight numeric inputs
   - Age slider
   - Gender selection
   
2. **Reactive Calculations**
   - BMI calculation based on user input
   - Health category determination
   - Personalized recommendations

3. **Dynamic Visualizations**
   - BMI gauge with health categories
   - Population comparison
   - Future BMI projections

## Visualizations

```{r, echo=FALSE, fig.height=4, fig.width=8}
# Create sample BMI gauge visualization
bmi <- 24.5  # Example BMI

# Create gauge plot
ggplot() +
  geom_rect(data = data.frame(
    xmin = c(0, 18.5, 25, 30),
    xmax = c(18.5, 25, 30, 40),
    ymin = c(-0.5, -0.5, -0.5, -0.5),
    ymax = c(0.5, 0.5, 0.5, 0.5),
    fill = c("#fff3cd", "#d4edda", "#f8d7da", "#f1b0b7")
  ), aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill), alpha = 0.7) +
  scale_fill_identity() +
  geom_text(data = data.frame(
    x = c(9.25, 21.75, 27.5, 35),
    y = c(0, 0, 0, 0),
    label = c("Underweight", "Normal", "Overweight", "Obese")
  ), aes(x = x, y = y, label = label), size = 4) +
  geom_segment(aes(x = bmi, y = -0.7, xend = bmi, yend = 0.7), 
              color = "#2c3e50", size = 1.5, arrow = arrow(length = unit(0.5, "cm"))) +
  geom_text(aes(x = bmi, y = -0.9, label = paste("Example BMI:", round(bmi, 1))),
           color = "#2c3e50", size = 5, fontface = "bold") +
  scale_x_continuous(limits = c(0, 40), breaks = c(0, 18.5, 25, 30, 40)) +
  scale_y_continuous(limits = c(-1, 1)) +
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  ) +
  labs(title = "BMI Scale")
```

## Try It Yourself

### Access the Application

- Visit: [https://github.com/galafis/BMI-Calculator-Shiny](https://github.com/galafis/BMI-Calculator-Shiny)
- Run the application locally or access it on shinyapps.io

### Features to Explore

- Calculate your own BMI
- View personalized health recommendations
- Explore different weight change scenarios
- Compare your BMI to population distributions

Thank you for your attention!
